---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r message=FALSE, warning=FALSE, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
library(ggplot2)
library(GEKS)
library(FEWS)
```

# GEKS

This package is a modification of the [FEWS package](https://github.com/MjStansfi/FEWS_package). Where the individual fixed effects calculation occurs, it has been replaced by a (modified to include imputed-tornqvist) GEKS calculation from [IndexNumR: A Package for Index Number Calculation](https://cran.r-project.org/web/packages/IndexNumR/vignettes/indexnumr.html#the-geks-method). 

* To see the original FEWS package written by Donal Lynch [go here](https://github.com/Donal-lynch/FEWS_package)

* To see the complete IndexNumR package written by Graham White [go here](https://github.com/grahamjwhite/IndexNumR)

From the IndexNumR package it uses:
* GEKS_w - modified
* GEKSIndex - modified
* utils
* fisher_t
* tornqvist_t
* ITRYGEKS_t - newly introduced

# Installation
GEKS is still in development. For now it can be installed from GitHub using the following code
```{r, eval = FALSE}

devtools::install_github("MjStansfi/GEKS_package")

#Other suggestions
# devtools::install_github("MjStansfi/FEWS_package")
# devtools::install_github("MjStansfi/TPDdecomp")
#devtools::install_github("grahamjwhite/IndexNumR")

# Once installed, the package can be loaded as usual
library(GEKS)
library(dplyr)
```

# Usage
The primary function provided by the GEKS package is the `GEKS()` function. Running `?GEKS()` should give all the required information on how using the function.
An example of running the `GEKS()` the function is shown below.


## Example

As part of the package, a couple of datasets are provided, including the Turvery dataset as found in the [Consumer Price Index Manual](https://www.ilo.org/wcmsp5/groups/public/---dgreports/---stat/documents/presentation/wcms_331153.pdf).

```{r data_viz}
ggplot(turvey, aes(x = month, y = price)) + 
  geom_line(aes(color = commodity)) + 
  geom_point(aes(color = commodity))+
  theme_bw() +
  ggtitle("Artificial prices of seasonal products Data created by R. Turvey")

```

The GEKS is calculated below with a mean splice and a window length of 13 months.

```{r fews_calculation, results='hide'}
turvey_GEKS <- GEKS(times = turvey$month,
                    price = turvey$price,
                    id = turvey$commodity,
                    window_length = 13,
                    weight = turvey$price * turvey$quantity,
                    splice_pos = "mean",
                    index_method = "tornqvist",
                    num_cores = NULL)

```

The resulting index is displayed below

```{r, geks_result}

ggplot(turvey_GEKS$geks, aes(x = price_date, y = fe_indexes)) + 
  geom_line() + 
  theme_bw() +
  ggtitle("GEKS with mean splice for Turvey data")+
  ylab("Price Index") + 
  xlab("Date")


```

We can compare the FEWS and GEKS as follows

```{r, geks_fews, results='hide'}
turvey_FEWS <-FEWS(times = turvey$month,
                    logprice = log(turvey$price),
                    id = turvey$commodity,
                    window_length = 13,
                    weight = turvey$price * turvey$quantity,
                    splice_pos = "mean",
                    num_cores = NULL)

```
The resulting index is displayed below
```{r, geks_fews_plot}
ggplot(turvey_FEWS$fews, aes(x = price_date, y = fe_indexes, colour = "FEWS")) +
  geom_line() +
  geom_line(aes(x =turvey_GEKS$geks$price_date, y= turvey_GEKS$geks$fe_indexes, colour = "GEKS-tornqvist"))+
  theme_bw() +
  ggtitle("Different multilateral, with mean splice for Turvey data")+
  ylab("Price Index") +
  xlab("Date")
```

# Imputation Törnqvist rolling year GEKS (ITRYGEKS)

Key functions in this package have been modified to calculate a Imputation Törnqvist rolling year GEKS (ITRYGEKS) index.

For this to work you need to employ the 'features' parameter of the GEKS function. Below we use this feature to calculate an index based on synthetic GfK data.

```{r, synthetic summary}
str(gfk_synthetic)
```

Now we wrangle this into the necessary format

```{r, format synthetic}

#Ensure that for each period there is only 1 observation of a given product
#ie sum all quantity and value for given month/product
gfk_synthetic <- gfk_synthetic%>%
  group_by(month_num,prodid_num)%>%
  mutate(quantity = sum(quantity),
         value = sum(value))%>%
  ungroup()%>%
  unique

#Calculate the unit value (price)
gfk_synthetic$uv <- gfk_synthetic$value/gfk_synthetic$quantity

#Extract data.frame containing features of interest
features <-gfk_synthetic[,grepl("char",colnames(gfk_synthetic))]
```

Once the dataframe is in the correct format we can run it through the main function with index_method set to 'impute-tornqvist'.

```{r, calc ITRYGEKS}
ITRYGEKS_index <- GEKS(times = gfk_synthetic$month_num,
                    price = gfk_synthetic$uv,
                    id = gfk_synthetic$prodid_num,
                    features = features,
                    window_length = 13,
                    weight = gfk_synthetic$value,
                    splice_pos = "mean",
                    index_method = "impute-tornqvist",
                    num_cores = NULL)


```

Lets run a standard tornqvist GEKS to compare 

```{r, calc GEKS}
GEKS_index <- GEKS(times = gfk_synthetic$month_num,
                    price = gfk_synthetic$uv,
                    id = gfk_synthetic$prodid_num,
                    features = NULL,
                    window_length = 13,
                    weight = gfk_synthetic$value,
                    splice_pos = "mean",
                    index_method = "tornqvist",
                    num_cores = NULL)


```


```{r, geks_geks_plot}
ggplot(ITRYGEKS_index$geks, aes(x = price_date, y = fe_indexes, colour = "ITRYGEKS")) +
  geom_line() +
  geom_line(aes(x =GEKS_index$geks$price_date, y= GEKS_index$geks$fe_indexes, colour = "GEKS-tornqvist"))+
  theme_bw() +
  ggtitle("Different multilateral, with mean splice for synthetic Gfk data")+
  ylab("Price Index") +
  xlab("Date")
```
